---
title: "Bartley Final Project"
format: html
editor: visual
---

#load in libraries

```{r}
library(tidyverse)
library(ggthemes)
library(tidycensus)
library(ggplot2)
```

#Load healthcare location data

```{r}
library(readr)
GeorgiaHealthCareFacilities <- read_csv("GeorgiaHealthCareFacilities.csv")
```

#Filter healthcare locations to only Fulton, DeKalb, Cobb, and Clayton counties (potential MARTA service area)

```{r}
atlanta_health <- GeorgiaHealthCareFacilities |>
  filter(County == "Fulton County" | County == "DeKalb County" | County == "Clayton County" | County == "Cobb County")
```

#Exported Atlanta Healthcare location data for use in QGIS

```{r}
#write.csv(atlanta_health, "atlanta_healthcare_data.csv")

```

#Finding the number of healthcare locations in each county that MARTA Services

```{r}
county_counts <- atlanta_health |>
  group_by(County) |>
  summarise(Healthcare_locations = n())
print(county_counts)
```

#Displaying the number of healthcare locations in each county on a graph

```{r}
ggplot(county_counts, aes(x = County, y = Healthcare_locations)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Healthcare Locations by County",
    x = "County",
    y = "Number of Healthcare Locations")
```

#Importing MARTA Train Station data

```{r}
MARTA_Rail_Stations <- read_csv("MARTA_Rail_Stations_-1706175048887107171.csv")
```

#Add a county column

```{r}
MARTA_Rail_Stations <- MARTA_Rail_Stations |>
  mutate(
    County = case_when(
      (City == "Atlanta" | City == "College Park" | City == "East Point" | City == "Sandy Springs") ~ "Fulton",
      (City == "Decatur" | City == "Brookhaven" | City == "Stone Mountain") ~ "DeKalb",
      (City == "Doraville") ~ "Clayton"
    )
  )
```

#Finding the number of MARTA stations in each county that MARTA Services

```{r}
MARTA_county_counts <- MARTA_Rail_Stations |>
  group_by(County) |>
  summarise(MARTA_Rail_Stations = n())
print(MARTA_county_counts)
```

#Chart of MARTA Rail Stations by County

```{r}
ggplot(MARTA_county_counts, aes(x = County, y = MARTA_Rail_Stations)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of MARTA Stations by County",
    x = "County",
    y = "Number of MARTA Stations")
```

#Importing healthcare data for locations within the one mile buffer

```{r}
MARTAroutes_noline <- read_csv("selectednoline.csv")
```

#Importing healthcare data for locations within the one mile buffer including the proposed new route

```{r}
MARTAroutes_newline <- read_csv("newlinehealth.csv")
```

#Counting rows in each data set and finding the difference

```{r}
oldroute_rows <- nrow(MARTAroutes_noline)
newline_rows <- nrow(MARTAroutes_newline)

row_difference <- abs(oldroute_rows - newline_rows)
percent_change <- abs(oldroute_rows - newline_rows) / ((oldroute_rows + newline_rows) / 2) * 100

row_difference
percent_change
```

There are 20 more healthcare locations included with the new route.

There is a 12.8% increase in healthcare locations within the new 1 mile boundary around MARTA train routes when the new route is included.
